Delocalization of Eigenvectors in Random Matrices

### 2025-03-22 Draft.tex
derivative

$$\Omega(t)_{xy}=G_{xy}(t)\left(\overline{\sum_j G_{xj}(t)S_{jy}(G_{yy}(t)-m(z))G_{jy}(t)}\right)+$$
$$+\overline{G_{xy}(t)}\left(\sum_j G_{xj}(t)S_{jy}(G_{yy}(t)-m(z))G_{jy}(t)\right)+\sum_{\alpha, \beta}|G_{xa}(t)|^2S_{\alpha, \beta}|G_{x\beta}(t)|^2$$
Hence, by Itô the $T$-equation becomes: 
\begin{equation*}
d{T} = {T}^2 dt + S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt \tag{3.1}
\end{equation*}
Now, define the \textit{time-dependent diffusion profile} as:$$\Theta_t = \frac{|m(z)|^2S}{1-t|m(z)|^2S}$$
Let us take its derivative w.r.t to $t$: 
$$\frac{d\Theta_t}{dt} = \frac{d}{dt}\frac{|m(z)|^2S}{1-t|m(z)|^2S} = \frac{-|m(z)|^2S(-|m(z)|^2S)}{(1-t|m(z)|^2S)^2} = \Theta_t^2 \Leftrightarrow d\Theta_t = \Theta_t^2 dt$$
Consider the fluctuation term $\mathcal{E}_t(z)$, i.e the difference between our $T$-matrix and the diffusion profile:
$$\E_t(z):T_t(z)-\Theta_t$$
We can compute its evolution equation $d\E_t(z) = dT_t(z)-d\Theta_t$ by substituting for what we found before:
$$\Rightarrow d\E_t(z) = T_t(z)^2dt - \Theta_t^2dt-S^{1/2}dM_t(z)S^{1/2}+S^{1/2}\Omega_t(z)S^{1/2}dt$$
But observe that:
$$T_t(z)^2-\Theta_t^2 =(\E_t(z)+\Theta_t(z))^2- \Theta_t(z)^2=\E_t(z)^2+\Theta_t\E_t(z)+\E_t(z)\Theta_t$$
\begin{equation*}\Rightarrow d\E_t(z)=\{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt +\E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2}
 \Omega_t(z) S^{1/2}dt\tag{3.2}\end{equation*}
Now, observe that the first term $\{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt$ makes this equation a matrix-valued linear SDE with a nonlinear stochastic term $W(t) = \E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt$. This means that we have to apply a method of variation of parameters (a matrix-valued Duhamel formula \cite{bandSDE}). For this, let us identify an integrating factor $U(t, s)$. Observe that: 
$$\partial_{t}\{\text{Id}+(t-s)\Theta_t\} = \Theta_t+(t-s)\Theta_t^2=\Theta_t\frac{1-t|m(z)|^2S+(t-s)|m(z)|^2S}{1-t|m(z)|^2S}=$$
$$=\Theta_t \frac{1-s|m(z)|^2S}{1-t|m(z)|^2S}=\Theta_t\{\text{Id}+(t-s)\Theta_t\}$$
This is equivalent to $\partial_t U(t, s) = \Theta_t U(t, s)$, which makes $U(t, s)$ our evolution operator, since also $U(t, t) = \text{Id}$. We can then define the integral form: $\E_t(z) = \int_0^t U(t, s) W(s) U(t, s)ds,$ and verify that it satisfies our SDE by using Leibniz's integral rule, we get:
$$d\E_t(z) = U(t, t) W(t) U(t, t)dt + \int_0^t d\left[U(t, s)W(s)U(t, s)\right]ds=$$
$$=W(t)dt + \int_0^t\left[\Theta_t U(t, s) W(s) U(t, s) + U(t, s)W(s)U(t, s)\Theta_t\right]dtds =$$
$$= \E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt + \{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt$$
Therefore, by the pathwise uniqueness of solutions to SDEs (Thm. 9.1 \cite{steele}), the following expression satisfies the same SDE:
\begin{equation*}\E_t(z)=\E_t^{D}(z)+\E_t^{M}(z)+\E_t^{S}(z)\tag{3.3}\end{equation*}
$$\E_t^{D}(z)=\int_0^t \{\text{Id} + (t-s)\Theta_t\} S^{1/2} \Omega(s) S^{1/2} \{\text{Id} + (t-s)\Theta_t\}ds,$$
$$\E_t^{M}(z)=-\int_0^t \{\text{Id} + (t-s)\Theta_t\}S^{1/2} dM(s)S^{1/2}\{\text{Id} +(t-s)\Theta_t\}ds,$$
$$\E_t^{S}(z)=\int_0^t \{\text{Id} + (t-s)\Theta_t\}\E_s^2(z) \{\text{Id} + (t-s)\Theta_t\}ds,$$

@@

\subsection{Quantum Diffusion}
\label{diffusion}
Recall our time-dependent $T$-equation $T(t)$ and diffusion profile $\Theta_t$. The standard $T$-matrix and diffusion profile $\Theta$ are just the latter two at the end of the spectral curve when $t=1$, i.e: $$T(z)_{ab} = T_{ab} = \sum_{x,y}S_{ax}^{\frac{1}{2}} |G_{xy}|^2S_{yb}^{\frac{1}{2}}, \quad \Theta := \frac{|m(z)|^2S}{1-|m(z)|^2S}$$
The first theorem established by Dubova and Yang (Theorem 2 \cite{bandSDE}) shows that in the bulk $T\approx \Theta$:
\begin{boxtheorem}{Theorem 2: (Quantum Diffusion)}
For $|E|<2$ fixed, assume $\exists \nu>0$, s.t. $\eta \asymp W^2 N^{-2}$ and $W\geq W^{8/11+\eta}$. Then: $$\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$
\end{boxtheorem}
This result has a physical interpretation. First, observe
that by \hyperref[2.1]{(2.1)} and  \hyperref[lemma-a11]{Lemma A.1.1}, $1-|m(z)|^2\asymp \eta\Leftrightarrow \exists \alpha>0, 1-|m(z)|^2\geq \alpha\eta$, we have that: 
 $$\Theta =\frac{|m(z)|^2S}{1-|m(z)|^2S} = \frac{|m(z)|^2S}{1-|m(z)|^2-|m(z)|^2(S-\text{Id})}\sim \frac{S}{\alpha\eta-(S-\text{Id})}$$
\textcolor{red}{Shouldn't this be a minus sign? Theorem 2 in \cite{bandSDE} states it with a plus sign.}
If we recall the phenomenology of our \hyperref[model]{model}, we can interpret the band width $W$ as the range of interactions in the particle system $\Z_N = \{1, \cdots, N\}$. In this context, since the doubly stochastic matrix is normalized as $\sum_x S_{xy}=1$ and $\sum_y S_{xy}=1$ for fixed $x, y\in \Gamma$, it has a unique interpretation as the transition matrix for a random walk on $\Z_N$. In this context, $S-\text{Id}$ is then clearly its generator, which allows us to consider $\Theta$ as its \textcolor{red}{resolvent?(neg. sign?)}. The spectral gap of this random walk on $\Gamma$ with steps of variance $W^2$ is of order $W^2N^{-2}$ (\textcolor{red}{reference?}). Standard bounds for diffusion \hyperref[lemma-a2]{Lemma A.2} imply that $\max_{a,b}\Theta_{ab} \lesssim W^{-1}\eta^{-1/2}$, whereas the latter theorem reveals a distinct scaling  $W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}\ll W^{-1}\eta^{-1/2}$. This in fact is precisely what characterized quantum diffusion, because instead of following standard resolvent bounds, the maximal entry of the diffusion profile $\Theta$ exhibits different scaling behavior, akin to the phenomenon of interference in quantum systems. Moreover, this quantum correction persists until the relaxation time $\eta^{-1}\asymp N^2W^{-2}$ (Thouless time \cite{17, 32, 33}), allowing us to prove delocalization (\hyperref[proof-deloc]{Theorem 2}).

@@

\begin{boxt}{Conditions}
\begin{enumerate}[label=(8.\arabic*)]
\item $\mathbb{E}\left[H_{ij}^\ell \left(\overline{H}_{ij}\right)^{k-\ell}\right] = \mathbb{E}\left[\left(H_{ij}^G\right)^\ell \left(\overline{H}_{ij}^G\right)^{k-\ell}\right], \quad 0\leq \ell\leq k\leq 5$
\item $\forall q\geq 1, \quad \mathbb{E}\left\vert H_{xy}\right\vert ^{2q}\prec W^{-q}$
\end{enumerate}
\end{boxt}
The first condition $(8.1)$ is the $\textit{Five Moment}$ matching assumption that gives the name of our result. The second $(8.2)$ is a moment decay condition that allows us to relax the Gaussian assumption, while still providing sufficient strength of our bounds. However, we still have to verify that each consecutive replacement is manageable. For this purpose, we will define the following interpolation for $t\in [0, 1]$:
$$tH^{(ij)} = H^{(i, j)}(H_{ij}, H_{ji}\mapsto t H_{ij}, tH_{ji})$$
$$tH^{(i, j), G} = H^{(i, j), G}(H^G_{ij}, H^G_{ij}\mapsto t H^G_{ij}, t H^{G}_{ij}).$$
Whenever the replacement step $(i, j)$ is implicit from context, we will drop the superscript. In order to make our theorem applicable to the context of delocalization, we want to consider the following function: 
$$F(H) = \left\vert
(H-z)_{ij}^{-1} - m(z)S_{ij}
\right\vert ^{2p} = \left\vert 
G_{ij}(z)-m(z)S_{ij}
\right\vert ^{2p} = |W_{ij}|^{2p}$$
Our reasoning is simple. Having familiarized ourselves with the flow method and the resolvent approach, it makes sense for us to choose a function that quantifies the deviation of resolvent entries form their deterministic approximations. By raising to power $2p$, we have better tools at our disposal, namely Chebyshev' and Markov's inequalities.

\noindent Consider the replacement at stage $ij$ for $t\in [0, 1]$:

$$F(tH_{ij}) = F(H_{11}^G, \cdots, H_{i, j-1}^G, tH_{ij}, H_{i, j+1}, \cdots H_{NN})$$
 By the chain rule: $$\partial_t F(tH_{ij}) = \partial_{H_{ij}} F(tH_{ij})H_{ij} + \partial_{H_{ji}} F(tH_{ij})\overline{H_{ij}}$$
(\textcolor{red}{prove by induction? - also justify the first formula} - Similarly, we have: 
$$\frac{d^k}{dt^k} F(tH_{ij}) = \sum_{k=0}^k c_{\ell, k}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell}$$

@@

>>>>

$$\Omega(t)_{xy}=G_{xy}(t)\left(\overline{\sum_j G_{xj}(t)S_{jy}(G_{yy}(t)-m(z))G_{jy}(t)}\right)+$$
$$+\overline{G_{xy}(t)}\left(\sum_j G_{xj}(t)S_{jy}(G_{yy}(t)-m(z))G_{jy}(t)\right)+\sum_{\alpha, \beta}|G_{xa}(t)|^2S_{\alpha, \beta}|G_{x\beta}(t)|^2$$
Hence, by Itô the $T$-equation becomes: 
\begin{equation*}
d{T} = {T}^2 dt + S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt \tag{3.1}
\end{equation*}
Now, define the \textit{time-dependent diffusion profile} as:$$\Theta_t = \frac{|m(z)|^2S}{1-t|m(z)|^2S}$$
Let us take its derivative w.r.t to $t$: 
$$\frac{d\Theta_t}{dt} = \frac{d}{dt}\frac{|m(z)|^2S}{1-t|m(z)|^2S} = \frac{-|m(z)|^2S(-|m(z)|^2S)}{(1-t|m(z)|^2S)^2} = \Theta_t^2 \Leftrightarrow d\Theta_t = \Theta_t^2 dt$$
Consider the fluctuation term $\mathcal{E}_t(z)$, i.e the difference between our $T$-matrix and the diffusion profile:
$$\E_t(z):T_t(z)-\Theta_t$$
We can compute its evolution equation $d\E_t(z) = dT_t(z)-d\Theta_t$ by substituting for what we found before:
$$\Rightarrow d\E_t(z) = T_t(z)^2dt - \Theta_t^2dt-S^{1/2}dM_t(z)S^{1/2}+S^{1/2}\Omega_t(z)S^{1/2}dt$$
But observe that:
$$T_t(z)^2-\Theta_t^2 =(\E_t(z)+\Theta_t(z))^2- \Theta_t(z)^2=\E_t(z)^2+\Theta_t\E_t(z)+\E_t(z)\Theta_t$$
\begin{equation*}\Rightarrow d\E_t(z)=\{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt +\E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2}
 \Omega_t(z) S^{1/2}dt\tag{3.2}\end{equation*}
Now, observe that the first term $\{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt$ makes this equation a matrix-valued linear SDE with a nonlinear stochastic term $W(t) = \E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt$. This means that we have to apply a method of variation of parameters (a matrix-valued Duhamel formula \cite{bandSDE}). For this, let us identify an integrating factor $U(t, s)$. Observe that: 
$$\partial_{t}\{\text{Id}+(t-s)\Theta_t\} = \Theta_t+(t-s)\Theta_t^2=\Theta_t\frac{1-t|m(z)|^2S+(t-s)|m(z)|^2S}{1-t|m(z)|^2S}=$$
$$=\Theta_t \frac{1-s|m(z)|^2S}{1-t|m(z)|^2S}=\Theta_t\{\text{Id}+(t-s)\Theta_t\}$$
This is equivalent to $\partial_t U(t, s) = \Theta_t U(t, s)$, which makes $U(t, s)$ our evolution operator, since also $U(t, t) = \text{Id}$. We can then define the integral form: $\E_t(z) = \int_0^t U(t, s) W(s) U(t, s)ds,$ and verify that it satisfies our SDE by using Leibniz's integral rule, we get:
$$d\E_t(z) = U(t, t) W(t) U(t, t)dt + \int_0^t d\left[U(t, s)W(s)U(t, s)\right]ds=$$
$$=W(t)dt + \int_0^t\left[\Theta_t U(t, s) W(s) U(t, s) + U(t, s)W(s)U(t, s)\Theta_t\right]dtds =$$
$$= \E_t^2(z)dt - S^{1/2} dM_t(z)S^{1/2} + S^{1/2} \Omega_t(z) S^{1/2}dt + \{\Theta_t \E_t(z)+\E_t(z)\Theta_t\}dt$$
Therefore, by the pathwise uniqueness of solutions to SDEs (Thm. 9.1 \cite{steele}), the following expression satisfies the same SDE:\textcolor{red}{Define:
\begin{equation*}\E_t(z)=\E_t^{D}(z)+\E_t^{M}(z)+\E_t^{S}(z)\tag{3.3}\end{equation*}
$$\E_t^{D}(z)=\int_0^t \{\text{Id} + (t-s)\Theta_t\} S^{1/2} \Omega(s) S^{1/2} \{\text{Id} + (t-s)\Theta_t\}ds,$$
$$\E_t^{M}(z)=-\int_0^t \{\text{Id} + (t-s)\Theta_t\}S^{1/2} dM(s)S^{1/2}\{\text{Id} +(t-s)\Theta_t\}ds,$$
$$\E_t^{S}(z)=\int_0^t \{\text{Id} + (t-s)\Theta_t\}\E_s^2(z) \{\text{Id} + (t-s)\Theta_t\}ds,$$}

@@

\subsection{Quantum Diffusion}
\label{diffusion}
Recall our time-dependent $T$-equation $T(t)$ and diffusion profile $\Theta_t$. The standard $T$-matrix and diffusion profile $\Theta$ are just the latter two at the end of the spectral curve when $t=1$, i.e: $$T(z)_{ab} = T_{ab} = \sum_{x,y}S_{ax}^{\frac{1}{2}} |G_{xy}|^2S_{yb}^{\frac{1}{2}}, \quad \Theta := \frac{|m(z)|^2S}{1-|m(z)|^2S}$$
The first theorem established by Dubova and Yang (Theorem 2 \cite{bandSDE}) shows that in the bulk $T\approx \Theta$:
\begin{boxtheorem}{Theorem 2: (Quantum Diffusion)}
For $|E|<2$ fixed, assume $\exists \nu>0$, s.t. $\eta \asymp W^2 N^{-2}$ and $W\geq W^{8/11+\eta}$. Then: $$\max_{x, y}|T_{xy}-\Theta_{xy}|\prec W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$$
\end{boxtheorem}
This result has a physical interpretation. First, observe
that by \hyperref[2.1]{(2.1)} and  \hyperref[lemma-a11]{Lemma A.1.1}, $1-|m(z)|^2\asymp \eta\Leftrightarrow \exists \alpha>0, 1-|m(z)|^2\geq \alpha\eta$, we have that: 
 $$\Theta =\frac{|m(z)|^2S}{1-|m(z)|^2S} = \frac{|m(z)|^2S}{1-|m(z)|^2-|m(z)|^2(S-\text{Id})}\sim \frac{S}{\alpha\eta-(S-\text{Id})}$$
If we recall the phenomenology of our \hyperref[model]{model}, we can interpret the band width $W$ as the range of interactions in the particle system $\Z_N = \{1, \cdots, N\}$. In this context, since the doubly stochastic matrix is normalized as $\sum_x S_{xy}=1$ and $\sum_y S_{xy}=1$ for fixed $x, y\in \Gamma$, it has a unique interpretation as the transition matrix for a random walk on $\Z_N$. In this context, $S-\text{Id}$ is then clearly its generator, which allows us to consider $\Theta$ as its resolvent. The spectral gap of this random walk on $\Gamma$ with steps of variance $W^2$ is of order $W^2N^{-2}$ \cite{spectralgap}. Standard bounds for diffusion \hyperref[lemma-a2]{Lemma A.2} imply that $\max_{a,b}\Theta_{ab} \lesssim W^{-1}\eta^{-1/2}$, whereas the latter theorem reveals a distinct scaling  $W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}\ll W^{-1}\eta^{-1/2}$. This in fact is precisely what characterized quantum diffusion, because instead of following standard resolvent bounds, the maximal entry of the diffusion profile $\Theta$ exhibits different scaling behavior, akin to the phenomenon of interference in quantum systems. Moreover, this quantum correction persists until the relaxation time $\eta^{-1}\asymp N^2W^{-2}$ (Thouless time \cite{17, 32, 33}), allowing us to prove delocalization (\hyperref[proof-deloc]{Theorem 2}).

@@

\begin{boxt}{Conditions}
\begin{enumerate}[label=(8.\arabic*)]
\item $\mathbb{E}\left[H_{ij}^\ell \left(\overline{H}_{ij}\right)^{k-\ell}\right] = \mathbb{E}\left[\left(H_{ij}^G\right)^\ell \left(\overline{H}_{ij}^G\right)^{k-\ell}\right], \quad 0\leq \ell\leq k\leq 5$
\item $\forall q\geq 1, \quad \mathbb{E}\left\vert H_{xy}\right\vert ^{2q}\prec W^{-q}$
\end{enumerate}
\end{boxt}
The first condition $(8.1)$ is the $\textit{Five Moment}$ matching assumption that gives the name of our result. The second $(8.2)$ is a moment decay condition that allows us to relax the Gaussian assumption, while still providing sufficient strength of our bounds. However, we still have to verify that each consecutive replacement is manageable. For this purpose, we will define the interpolation for $t\in [0, 1]$ at entries $(ij)$ and $(ji)$, where with slight abuse of notation, we let:
$$tH_{ij} = H^{(i, j)}[H_{ij}, H_{ji}\mapsto t H_{ij}, tH_{ji}]$$
$$tH_{ij}^{G} = H^{(i, j),G}[H^G_{ij}, H^G_{ij}\mapsto t H^G_{ij}, t H^{G}_{ij}].$$
Whenever the replacement step $(i, j)$ is implicit from context, we will drop the superscript. In order to make our theorem applicable to the context of delocalization, we want to consider the following function: 
$$F(H) = \left\vert
(H-z)_{ij}^{-1} - m(z)S_{ij}
\right\vert ^{2p} = \left\vert 
G_{ij}(z)-m(z)S_{ij}
\right\vert ^{2p} = |W_{ij}|^{2p}$$
Our reasoning is simple. Having familiarized ourselves with the flow method and the resolvent approach, it makes sense for us to choose a function that quantifies the deviation of resolvent entries form their deterministic approximations. By raising to power $2p$, we have better tools at our disposal, namely Chebyshev' and Markov's inequalities. Let us apply our strategy to $F$ at step $ij$ for $t\in [0, 1]$:

$$F(tH_{ij}) = F(H_{11}, \cdots, H_{i, j-1}, tH_{ij}, H_{i, j+1}^G, \cdots H_{NN}^G)$$
In order to apply the Taylor series for the error terms, let us calculate the derivatives: $$\partial_t F(tH_{ij}) = \partial_{H_{ij}} F(tH_{ij})H_{ij} + \partial_{H_{ji}} F(tH_{ij})\overline{H_{ij}}$$
We claim that the following represents the $k$th derivative: 
$$\frac{d^k}{dt^k} F(tH_{ij}) = \sum_{\ell=0}^k \binom{k}{\ell}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell}$$
Assume the latter for an induction hypothesis, having already verified the base case. Then: 
$$\frac{d^{k+1}}{dt^{k+1}} F(tH_{ij}) = \frac{d}{dt}\left[\frac{d^k}{dt^k} F(tH_{ij}) \right]=\frac{d}{dt}\left[ \sum_{\ell=0}^k \binom{k}{\ell}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell}\right]=$$
$$=\sum_{\ell=0}^k\binom{k}{\ell}\frac{d}{dt}\left[\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell} F(tH_{ij})\right]H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell}$$

\noindent By the chain rule for any $\ell$ and $m$:
$$\frac{d}{dt}\left[\partial_{H_{ij}}^\ell \partial_{H_{ji}}^m F(tH_{ij})\right] = \partial_{H_{ij}}^{\ell+1} \partial_{H_{ji}}^m F(tH_{ij})H_{ij} + \partial_{H_{ij}}^\ell \partial_{H_{ji}}^{m+1} F(tH_{ij})\overline{H_{ij}}$$
Substituting in the derivative formula gives us:
$$
= \sum_{\ell=0}^k \binom{k}{\ell}\left[\partial_{H_{ij}}^{\ell+1} \partial_{H_{ji}}^{k-\ell} F(tH_{ij})H_{ij} + \partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell+1} F(tH_{ij})\overline{H_{ij}}\right]H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell}$$
$$
= \sum_{\ell=0}^k \binom{k}{\ell}\partial_{H_{ij}}^{\ell+1} \partial_{H_{ji}}^{k-\ell} F(tH_{ij})H_{ij}^{\ell+1} \left(\overline{H_{ij}}\right)^{k-\ell} + \sum_{\ell=0}^k \binom{k}{\ell}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell+1} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell+1}
$$
$$
= \sum_{\ell=1}^{k+1} \binom{k}{\ell-1}\partial_{H_{ij}}^{\ell} \partial_{H_{ji}}^{k+1-\ell} F(tH_{ij})H_{ij}^{k+1-\ell} \left(\overline{H_{ij}}\right)^{k+1-\ell} + \sum_{\ell=0}^k \binom{k}{\ell}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k-\ell+1} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k-\ell+1}
$$
But observe that we can split the sums and group it as follows:
$$=\binom{k}{0}\partial_{H_{ij}}^0\partial_{H_{ji}}^{k+1} F(tH_{ij})\left(\overline{H_{ij}}\right)^{k+1}+\binom{k}{k}\partial_{H_{ij}}^{k+1}\partial_{H_{ji}}^{0} F(tH_{ij})\left(\overline{H_{ij}}\right)^{k+1}+$$
$$=+\sum_{\ell=1}^k \left(\binom{k}{\ell-1}+\binom{k}{\ell}\right)\partial_{H_{ij}}^{\ell} \partial_{H_{ji}}^{k+1-\ell} F(tH_{ij})H_{ij}^{\ell} \left(\overline{H_{ij}}\right)^{k+1-\ell} $$
But since $\binom{k}{k} = \binom{k+1}{k+1}$ and $\binom{k}{\ell-1}+\binom{k}{\ell} = \binom{k+1}{\ell}$, this completes the induction:
$$\frac{d^{k+1}}{dt^{k+1}} F(tH_{ij}) = \sum_{\ell=0}^{k+1} \binom{k+1}{\ell}\partial_{H_{ij}}^\ell \partial_{H_{ji}}^{k+1-\ell} F(tH_{ij})H_{ij}^\ell \left(\overline{H_{ij}}\right)^{k+1-\ell}$$

\subsection*{Taylor series}

@@

%\bibitem{spectralgap} Find reference
